---
title: "Capstone - Milestone report"
author: "Stavros Oikonomou"
date: "3/6/2020"
output: html_document
---
### Summary

The goal of this project is just to display that you've gotten used to working with the data and that you are on track to create your prediction algorithm.

This is a milestone report for peer graded assigment as part of Data Science Captone Course from Coursera in Week 2. The objective of this document is as follows,

1. Demonstrate that you’ve downloaded the data and have successfully loaded it in.
2. Create a basic report of summary statistics about the data sets.
3. Report any interesting findings that you amassed so far.
4. Get feedback on your plans for creating a prediction algorithm and Shiny app.

This report also will be served as a base for creating the next assignment report, hence it should be as clear and concise as possible. The content of this report will be structured to 5 sections as per the objective mentioned above.


#### Loading the libraries
```{r,warning=FALSE,message=FALSE}
library(ggplot2)
library(tm)
library(stringi)
library(textmineR)
library(NLP)
```

#### Downloading the dataset
```{r}
if (!file.exists("Coursera-SwiftKey.zip")){
  download.file("https://d396qusza40orc.cloudfront.net/dsscapstone/dataset/Coursera-SwiftKey.zip","Coursera-SwiftKey.zip",
                method = "auto")
}
```

Unzipping the zip file
```{r}
blog_file<- "final/en_US/en_US.blogs.txt"
twit_file <- "final/en_US/en_US.twitter.txt"
news_file <- "final/en_US/en_US.news.txt"

if (!file.exists(blog_file) || !file.exists(twit_file) || !file.exists(news_file) ){
  unzip("Coursera-SwiftKey.zip")}
```

#### Loading the text files
We gonna work with english dataset. There are three files for blogs twitter and news
```{r,warning=FALSE,message=FALSE}
blogs   <- readLines(blog_file, encoding="UTF-8")
twitter <- readLines(twit_file, encoding="UTF-8")
news    <- readLines(news_file, encoding="UTF-8")
```

#### Creating the summary
The basic details for the three datasets. The number of words, the number of lines, the number of characters and the size of files.
```{r}
summary <- data.frame('File Name ' = c("Blogs","News","Twitter"),
                      'words_counts'= sapply(list(blogs, news, twitter), function(x) {sum(stri_count_words(x))}),
                      'line_counts' = sapply(list(blogs, news, twitter), stri_stats_general)[1,],
                      'Chars'= sapply(list(blogs, news, twitter), stri_stats_general)[3,],
                      'Size'= sapply(list(blogs, news, twitter),function(x){format(object.size(x),"MB")}))
summary
```

#### Data sample
Because the dataset is large we gonna create a subset with 1% of the data per file
```{r}
set.seed(1234)
sample_size <- 0.01

data_sample <- c(sample(blogs, length(blogs) * sample_size),
                 sample(news, length(news) * sample_size),
                 sample(twitter, length(twitter) * sample_size))
```

#### Building the Corpus
```{r}
corpus <- VCorpus(VectorSource(data_sample))
```

#### Data cleaning and preprocessing
We gonna:

* Turn every letter to lowercase
* Removing double spaces
* Removing punctuations
* Removing every number
* Removing standar words we dont need
```{r}
corpus<-tm_map(corpus, content_transformer(tolower))
corpus<-tm_map(corpus, stripWhitespace)
corpus<-tm_map(corpus, removePunctuation)
corpus<-tm_map(corpus, removeNumbers)
corpus<-tm_map(corpus, removeWords, stopwords('english'))
```

#### Creating N-grams for our data
We have cleaned and sampled our data. we have done some preprocessing for our data. Now we can build our basic unigram, bigram and trigram. We choose 99.99% sparsity in the range from bigger zero to smaller one, so we gonna keep 99.99% of the most frequent words.
```{r}
BigramTokenizer <-function(x) unlist(lapply(ngrams(words(x), 2), paste, collapse = " "), use.names = FALSE)
TrigramTokenizer <- function(x) unlist(lapply(ngrams(words(x), 3), paste, collapse = " "), use.names = FALSE)

freq_df <- function(tdm){
  freq <- sort(rowSums(as.matrix(tdm)), decreasing=TRUE)
  freq_df <- data.frame(word=names(freq), freq=freq)
  return(freq_df)
}

unigram <- removeSparseTerms(TermDocumentMatrix(corpus), 0.9999)
unigram_freq <- freq_df(unigram)

bigram <- removeSparseTerms(TermDocumentMatrix(corpus, control = list(tokenize = BigramTokenizer)), 0.9999)
bigram_freq <- freq_df(bigram)

trigram <- removeSparseTerms(TermDocumentMatrix(corpus, control = list(tokenize = TrigramTokenizer)), 0.9999)
trigram_freq <- freq_df(trigram)
```

#### Plotting the most common unigrams,bigrams and trigrams
```{r}
ggplot(unigram_freq[1:10,], aes(reorder(word,-freq),freq, fill=word)) + geom_bar(stat="identity") +
  labs(x="words", y="Frequency") + ggtitle("10 most common Unigrams") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, size = 12, hjust = 1)) 
```

```{r}
ggplot(bigram_freq[1:10,], aes(reorder(word,-freq),freq, fill=word)) + geom_bar(stat="identity") +
  labs(x="words", y="Frequency") + ggtitle("10 most common Bigrams") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, size = 12, hjust = 1)) 
```

```{r}
ggplot(trigram_freq[1:10,], aes(reorder(word,-freq),freq, fill=word)) + geom_bar(stat="identity") +
  labs(x="words", y="Frequency") + ggtitle("10 most common Trigrams") + 
  theme(legend.position = "none", axis.text.x = element_text(angle = 45, size = 12, hjust = 1))
```

### Conclusion and Next Steps
We just finish the initial exploratory analysis. The next will be to build a predictive model that gonna help us predict the next word using the most frequent options. he algorithm will then be deployed in a Shiny app and will suggest the most likely next word after a phrase is typed.









